\documentclass[12pt]{article}

\usepackage{libertine}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{microtype}
\usepackage[english, russian]{babel}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{geometry}
\usepackage{fancyhdr}
\usepackage{natbib}
\usepackage{xcolor}
\usepackage{titlesec}
\usepackage{booktabs}
\usepackage[colorlinks=true,linkcolor=black,citecolor=black,urlcolor=black]{hyperref}

\geometry{
    margin=1in,
    headheight=12pt
}
\setcitestyle{square}

\pagestyle{fancy}
\fancyhf{}
\renewcommand{\headrulewidth}{0.4pt}
\renewcommand{\footrulewidth}{0.4pt}
\fancyhead[L]{\small Актуальные вопросы междисциплинарных научных исследований}
\fancyhead[R]{\small\thepage}
\fancyfoot[C]{\small Караганда, 2025}

\titleformat{\section}
    {\normalfont\large\bfseries}{\thesection}{1em}{}
\titlespacing*{\section}
    {0pt}{2.5ex plus 1ex minus .2ex}{1.5ex plus .2ex}

\makeatletter
\renewcommand{\maketitle}{%
    \begin{center}
        \Large\@title
        
        \vspace{0.4cm}
        \large\@author
        
        \vspace{0.5cm}
        \normalsize\textit{\textsuperscript{1}L.N. Gumilyov Eurasian National University, Astana, Kazakhstan\\
          E-mail: bokayev2011@yandex.ru\\
          \textsuperscript{2}Institute of Mathematical of the  Czech Academy of Sciences\\
          E-mail: gogatish@math.cas.cz\\
          \textsuperscript{3}L.N. Gumilyov Eurasian National University, Astana, Kazakhstan\\
          E-mail: nurgul.kuzeubaeva@mail.ru\\}
          \textsuperscript{4}Department of Mathematics, Kirikkale University, Kirikkale, T\" urkiye\\
          E-mail: tugceunver@kku.edu.tr\\
    \end{center}
}
\makeatother

% Paper info
\title{Reduction theorems for the discrete nonlinear operator on the cones of monotone sequences}

% Uncomment and update the following line upon acceptance
% \author{Author Name(s)\\ % Institution Name(s)}

%Comment the following line upon acceptance
\author{ N.~Bokayev.\textsuperscript{1}, A.~Gogatishvili.\textsuperscript{2}, N.~Kuzeubayeva\textsuperscript{3} , N.~"Unver\textsuperscript{4}}


\date{}

\begin{document}

\maketitle


Let $\{u_k\}$,  $\{a_n\}$ and $\{b_n\}$ be given
non-negative sequences. Let $p, q\in (0,\infty)$. We will investigate the following inequalities:

\begin{equation}  \label{T1}
\left(\sum_{n=1}^{\infty}
\left(\sup_{n\le i<\infty} u_i\sum^i_{k=1}x_k\right)^q a_n\right)^{\frac{1}{q}} \le C
\left(\sum^\infty_{n=1} x_n ^p b_n\right)^{\frac{1}{p}},
\end{equation}   
\begin{equation}  \label{T2}
\left(\sum_{n=1}^{\infty}
\left(\sup_{1\le i<n} u_i\sum^i_{k=1}x_k\right)^q a_n\right)^{\frac{1}{q}} \le C
\left(\sum^\infty_{n=1} x_n ^p b_n\right)^{\frac{1}{p}}
\end{equation}
for non-negative, non-increasing sequences $x = \{x_n\}$   and the
constant $C > 0$  is independent of $x$. 

\textbf{Theorem 1.}  \label{Eg_1} Let $0<q\leq\infty,\,  1<p<\infty$. Assume that $\{a_n\}$ and $\{b_n\}$ are given
non-negative weight sequences. %and put $B_n:=\sum_{i=1}^n b_i$  and  $\sum_{n=1}^\infty b_n=\infty$.  
Then the inequality \eqref{T1} holds  for all non-negative, non-increasing sequences $\{x_n\}$ if and only if the following  inequality:
\begin{equation}
    \label{T11}
\left(\sum_{n=1}^\infty\left(\sup_{n\le i<\infty} u_i\sum_{k=1}^i\left(\sum_{j=k}^\infty y_j\right)\right)^q a_n\right)^{\frac{1}{q}}\leq C\left(\sum_{n=1}^\infty y_n^pB_{n+1}^{p-1}B_n b_{n+1}^{1-p}\right)^{\frac{1}{p}}, 
\end{equation}
holds for all non-negative sequences $\{y_n\}$, where  $B_n =\sum^n_{k=1}b_k, \quad n\in {\mathbb N}$.

\begin{theorem} \label{Th4}
\textbf{Theorem 2.}  \label{Eg_1}
Let $0<q\leq\infty,\,  1<p<\infty$. Assume that $\{a_n\}$ and $\{b_n\}$ are given
non-negative weight sequences.% and put $B_n:=\sum_{i=1}^n b_i$.  
Then the inequality \eqref{T2} holds  for all non-negative, non-increasing sequences $\{x_n\}$ if and only if for any $\alpha>0$ the following inequality:
\begin{equation}
    \label{T41}
\left(\sum_{n=1}^\infty\left(\sup_{1\le i<n } 
u_i\sum_{i=n}^\infty \frac{1}{B_i^{\alpha+1}}\left(\sum_{k=1}^i  B_k^{\alpha+1}y_k\right)\right)^q a_n\right)^{\frac{1}{q}}\leq C\left(\sum_{n=1}^\infty y_n^pB_{n}^{p} b_{n}^{1-p}\right)^{\frac{1}{p}}, 
\end{equation}
holds for all non-negative sequences $\{y_n\}$, where  $B_n =\sum^n_{k=1}b_k, \quad n\in {\mathbb N}$.

\end{theorem}

In the continuous case, similar questions were considered in [1] and [2].

\begin{thebibliography}{00}\label{ref:ref}

\bibitem{1}
A.Gogatishvili, L. Pick and T. \"Unver, "Weighted inequalities for discrete iterated kernel operators", Math. Nachr. 295:11 (2022), 2171-2196.

\bibitem{2}
A.Gogatishvili and V.D.Stepanov., "Reduction theorems for weighted integral inequalities on the cone of monotone functions", Uspekhi Mat. Nauk., 405:1 (2013), 156-172.

\end{thebibliography}
\end{document}

